---
title: "Math 245 HW 4"
author: 'Name: Noam Benkler '
date: 'I worked with: Christian Zaytoun'
output:
  github_document: default
---
```{r, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment=NULL)

library(Sleuth3)
library(ggformula)
library(dplyr)
library(ggplot2)
library(stargazer)
library(broom)
library(tidyr)
library(GGally)
library(car)
library(gmodels)

data(ex0914)
?ex0914
data(case1002)
?case1002
```

## Problem 1 
### (a) 

```{r}
    scatterplotMatrix(~  Talk + Bank + Walk + Heart, data = ex0914)
```

### (b)

```{r}
hearts_lm <- lm(Heart ~ Bank + Walk + Talk, data = ex0914)
tidy(hearts_lm)
```

### (c)
it does not appear that that variance of the residuals increases as the fitted values increase
```{r}
hearts.re <- augment(hearts_lm)
gf_point(.resid ~ .fitted, data = hearts.re) %>% gf_hline(yintercept = 0)
```

### (d)
$\widehat{\mu}(Heart | Bank) = \underset{(6.34)}{3.18} + \underset{(0.20)}{0.41 (Bank)}$
All else held constant, for every point increase in mean time it takes bank clerks to make change for $20 average age adjusted death rates increase by 0.41 points.
$\widehat{\mu}(Heart | Walk) = \underset{(6.34)}{3.18} + \underset{(0.20)}{0.45 Walk}$
All else held constant, for every point increase in mean pedestrian walking speed average age adjusted death rates increase by 0.45 points.
$\widehat{\mu}(Heart | Talk) = \underset{(6.34)}{3.18} - \underset{(0.22)}{0.18 (Talk)}$
All else held constant, for every point increase in mean postal clerk talking speed average age adjusted death rates decrease by 0.18 points.

## Problem 2

### (a)
$\widehat{\mu}(Depression | Age, Education ) = {\widehat{\beta}_0} + {\widehat{\beta}_1 Age} + {\widehat{\beta}_2 Education,i} + {\widehat{\beta}_3 Education,ii} +{\widehat{\beta}_4 Education,iii}$
$ parameter = \beta_3-\beta_2$

### (b)
$\widehat{\mu}(Depression | Age, Education ) = {\widehat{\beta}_0} + {\widehat{\beta}_1 Age} + {\widehat{\beta}_2 Education,i} + {\widehat{\beta}_2 Education,ii} +{\widehat{\beta}_3 Education,iii}$
$ parameter = \beta_3-\beta_2$

### (c)
$\widehat{\mu}(Depression | Age, Education ) = {\widehat{\beta}_0} + {\widehat{\beta}_1 (Age-k)^2} + {\widehat{\beta}_2 Education,i} + {\widehat{\beta}_3 Education,ii} +{\widehat{\beta}_4 Education,iii}$
where k = some constant that stands for the mean age at which the lowest point in mean depression is reached


## Problem 3

### (a)
  d.f. for $\sigma$ estimation = 32

### (b)
$H_o: \beta_4 = \beta_5$
$H_a: \beta_4 \neq \beta_5$
According to diagram 10.18 the p-value for $\beta_4 = \beta_5$ is 0.0083 which is enough for us to reject the null hypothysis that species two has the same regression as species 1

### (c)
With 95% confidence the slope for species 3 exceeds the slope for species 1 by between 0.693 and -6.481 points.

```{r}
SE = 1.761
slope = -2.486 - 0.4083 #slope 3 - slope 1
t.critical <- qt(.025, df = 32)
slope + t.critical*SE
slope - t.critical*SE
```



## Problem 4

$ F = \frac{MSModel}{MSResid} =  \frac{\frac{2.384}{2}}{0.18741}=6.362$
p-value = 0.8596
```{r}
pf(6.362, 4,2)
```

## Problem 5

### (a)
```{r}
case.lm <- lm(log(Energy) ~ log(Mass) + Type, data = case1002)
stargazer(case.lm, type = "text")
```
$\widehat{\mu}(log(Energy) | log(Mass), Species) = \underset{(0.150)}{-1.498} + \underset{(0.045)}{0.815(bats:e)}$

$\widehat{\mu}(log(Energy) | log(Mass), Species) = \underset{(0.150)}{-1.498} - \underset{(0.203)}{0.079(bats:non-e)}$

$\widehat{\mu}(log(Energy) | log(Mass), Species) = \underset{(0.150)}{-1.498} + \underset{(0.158)}{0.024(birds:non-e)}$


### (b)
$\widehat{\mu}(log(Energy) | log(Mass), Species ) = {\widehat{\beta}_0} + {\widehat{\beta}_1 log(Mass)} + {\widehat{\beta}_2 (Bats:non-e)} + {\widehat{\beta}_3 (Birds:non-e)} +{\widehat{\beta}_4 (Bats:e)}$
$H_0: \beta_3 = \beta_4 $
$H_a: \beta_3 \neq \beta_4 $
There is sufficient evidence to reject the $H_0$ that $\beta_3 = \beta_4$.
```{r}
estimable(case.lm, c(0, 1, -1, 0))
```


### (c)
$\widehat{\mu}(log(Energy) | log(Mass), Species ) = {\widehat{\beta}_0} + {\widehat{\beta}_1 log(Mass)} + {\widehat{\beta}_2 (Bats:non-e)} + {\widehat{\beta}_3 (Birds:non-e)} +{\widehat{\beta}_4 (Bats:e)}$
$H_0: \beta_2 = \beta_3 $
$H_a: \beta_2 \neq \beta_3 $
```{r}
estimable(case.lm, c(0, 0, -1, 1))
```
Comparing parameters, we observe a p-value of 0.384. Therefore we can not reject the null hypothesis that non-echolocating bats have different log energy outputs than non-
echolocating birds.

### (d)
```{r}

case_lm <- augment(case.lm)
gf_point(log(Energy) ~ log(Mass), data = case1002, color = ~Type, shape ~Type) %>%
gf_lm(.fitted ~ log.Mass., data= case_lm) %>% gf_refine(scale_color_discrete()) %>%
  gf_labs(x="Mass(g)", y = "Energy(kJ)")
```

